Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-26 paper/high-assurance-contracts.tex
supposedly well-formed Fib example was not; fixed

\subsubsection*{Wellformedness}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages being targeted by SPLAT do not directly support streams.  We
are currently developing a means of translating away such deep
temporal accesses by the introduction of intermediate variables. For
now, we focus on the notion of \emph{wellformed} equations that needs
to hold. The guiding intuition is that a wellformed {\Eqs} allows
access at most one step in the past, and accesses of a past value for
a variable are only possible until the variable is updated. In
informal terms, this means that ${\Eqs}\cdot G$ in a code contract
must be such that:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-wellformed manner, by
{\small
\begin{verbatim}
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{verbatim}
}
\noindent or by (again non-wellformed)
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib)).
\end{verbatim}
}
\noindent or by (wellformed):
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  A = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(A)
\end{verbatim}
}
A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

>>>>

\subsubsection*{Wellformedness}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages being targeted by SPLAT do not directly support streams.  We
are currently developing a means of translating away such deep
temporal accesses by the introduction of intermediate variables. For
now, we focus on the notion of \emph{wellformed} equations that needs
to hold. The guiding intuition is that a wellformed {\Eqs} allows
access at most one step in the past, and accesses of a past value for
a variable are only possible until the variable is updated. In
informal terms, this means that ${\Eqs}\cdot G$ in a code contract
must be such that:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-wellformed manner, by
{\small
\begin{verbatim}
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{verbatim}
}
\noindent or by (again non-wellformed)
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
\end{verbatim}
}
\noindent or by (wellformed):
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  F2 = 42 -> pre(F1)
  F1 = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else F1 + F2
\end{verbatim}
}
A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.
